---
title: Generating Dynamic Contents in R Markdown and Quarto
slug: dynamic-rmd-quarto
date: '2023-01-22'
description: |
    Automate report generation in Quarto with parameterized documents.
tags:
    - Quarto
components:
    - iframe
---




A common scenario in my day-to-day data analysis is that I have many objects, typically data frames, and for each of them I want to create a new section in my Quarto/R Markdown document with its summary statistics. The operations are usually the same; only the data is different. The manual way is to copy and paste the same chunk of code for generating the summaries and change only the variable name. For instance:



````{verbatim}

## Section for Data A

Data A has `r nrow(data_a)` rows and `r ncol(data_a)` columns.


```{r}
summary_fn(data_a)
```

## Section for Data B

Data B has `r nrow(data_b)` rows and `r ncol(data_b)` columns.


```{r}
summary_fn(data_b)
```

## Section for Data C

...
````

This becomes a pain when there are many objects and I have to change the names one by one. This post shows how to automate this task with `knit_child` from the `knitr` rendering engine.


## Using the `knitr` Engine

The `knitr` engine has built-in support for dynamic, parameterized documents. For R Markdown, you can pass data to the document using the `params` argument in `rmarkdown::render()` (learn more at <https://bookdown.org/yihui/rmarkdown/parameterized-reports.html>). This is intended for rendering one big parameterized `.Rmd` document.

To insert sub-contents into either `.qmd` or `.Rmd` files, `knitr::knit_child` is the function you need. Similar to `params` in `rmarkdown::render`, you can pass an environment object to `knitr::knit_child` using the `envir` argument. `knit_child` returns a character string that can be embedded directly into a document, so it's used together with the chunk option `output = 'asis'` (Quarto) or `results = 'asis'` (R Markdown).

`knit_child` accepts either an inline string `text` or a file path `file`. Below we use the inline string to generate a summary of the `iris` data frame.




```{r}
#| echo: fenced
#| output: asis
# chunk option output = 'asis'
res <- knitr::knit_child(text = c(
    "```{r}",
    "summary(data)",
    "```"
), envir = rlang::env(data = iris), quiet = TRUE)
cat(res, sep = "\n")
```


Putting it all together, we can divide the job into a main document `main.Rmd` and a template document `_template.Rmd`. In the main document, we collect all the data, loop through them and insert them into the template document using `knitr::knit_child`. Then we can use the `asis` output option to include the template document in the main document.

::: {.column-margin}

This will also work for Quarto documents as long as we are using the knitr engine.

:::

The example main and template documents are shown below:


:::: {.two-column .wider}

::: {.col-span-1}

In `main.Rmd`

```{r}
#| collapse: false
#| comment: ""
#| class-output: "markdown"
#| echo: false
x <- xfun::read_utf8("_main.Rmd")
cat(x, sep = "\n")
```

:::

::: {.col-span-1}

In `_template.Rmd`

```{r}
#| collapse: false
#| comment: ""
#| class-output: "markdown"
#| echo: false
x <- xfun::read_utf8("_template.Rmd")
cat(x, sep = "\n")
```

:::

::::

I've defined a wrapper function `render_child()` that uses `knit_child()` under the hood. The key is that you can pass arbitrary values using the `envir` argument: pass whatever the template document needs as a named list and convert it into an environment object with `rlang::env()`.

When you run the main document, it generates a list of random data frames, generates a child document for each using the template document, and inserts the results back into the main document. The result is shown below:

::: {.column-body-outset }

<iframe src = "https://bookdown.org/qiushi/dynamic_contents_in_r_markdown_with_knit_child/" > </iframe>

:::



### A Note on `knitr::knit_expand`


[R Markdown Cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/knit-expand.html) also introduces `knitr::knit_expand()`, which inserts contents into a template. For example:

```{r}
knitr::knit_expand(
    text = "The value of `a` is {{a}}",
    a = 1
)
```

Since you can also pass a `file` argument to `knit_expand`, I was tempted to use this function with `knit_child` when I started writing this post. For example:

```{r}
#| eval: false
res <- knitr::knit_expand(
    file = "template.Rmd",
    data = iris[1:5, ]
)
```

Then you can access `iris` in the template document like so:

````{verbatim, filename="template.Rmd"}
---
title: template document
---

```{r}
{{ data }}
```
````

But this will not work. `knitr::knit_expand` simply finds all the placeholders marked by `{{ }}`, interpolates them and then passes the text to the knitting functions. This works fine if you only need to insert simple primitives, such as strings and numbers, but not if you want to pass data.frames or other complex objects. The knitting functions will then see a code chunk like this:

````{verbatim, filename="template.Rmd"}
```{r}
c(5.1, 4.9, 4.7, 4.6, 5)
c(3.5, 3, 3.2, 3.1, 3.6)
c(1.4, 1.4, 1.3, 1.5, 1.4)
c(0.2, 0.2, 0.2, 0.2, 0.2)
c(1, 1, 1, 1, 1)
```
````

which is not valid R syntax.

## Without `knitr`

If you are using Quarto with an engine other than `knitr`, I found no construct similar to `knitr::knit_child`. Although Quarto offers a native [variables](https://quarto.org/docs/authoring/variables.html) mechanism for reading values from configuration files, it's intended for static data that should be shared across multiple documents. It also has the same problem as `knitr::knit_expand`: you are limited to the data structures YAML supports.
